Finiteness in Cubical Type Theory

### 2020-06-17 sections/finiteness-predicates.tex
more introduction
moved definitions to their place of first usage

  Moreover, the inductive type and the container-based type are equivalent, as
  is proven in our formalisation.
\end{definition}
\begin{definition}[Containers] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a pair
  \(S , P\) where \(S\) is a type, the elements of which are called
  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
  the elements of \(P(s)\) are called the \emph{positions} of a container.
  We ``interpret'' a container into a functor defined like so:
  \begin{equation} \label{container-interp}
    \llbracket S , P \rrbracket(A) \coloneqq \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
  \end{equation}
  Membership of a container can be defined like so:
  \begin{equation} \label{container-membership}
    x \in \mathit{xs} \coloneqq \text{fib}_{\text{snd}(\mathit{xs})}(x)
  \end{equation}
\end{definition}
\begin{definition}[\(\mathbf{Fin}\)] \label{Fin}
  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
  define it the standard way:
  \begin{equation}
    \begin{alignedat}{3}
      & \mathbf{Fin}(0)     && \coloneqq \bot ; \\
      & \mathbf{Fin}(n + 1) && \coloneqq \top \uplus \mathbf{Fin}(n) ;
    \end{alignedat}
  \end{equation} 
\end{definition}
\subsection{Instances}
Now that we have a suitable definition of finiteness, we will next prove that
some things are finite.
\begin{lemma}
  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
\end{lemma}
\begin{proof}
  These three types are represented by fairly simple finite sets:
  \begin{align}
    \bot &= \left\{  \right\} \\
    \top &= \left\{ \text{tt}  \right\} \\
    \mathbf{Bool} &= \left\{  \text{false}, \text{true} \right\}
  \end{align}
  Which leads to quite simple proofs of finiteness.

@@

First, we will need to define our terms: in HoTT, surjections are a little more
complex than what you'd find in either MLTT or classical mathematics.
\begin{definition}[Surjections] \label{surjections}
  We define both surjections and \emph{split} surjections here \cite[definition
  4.6.1]{hottbook}.
  \begin{alignat}{3}
    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \label{surj-eqn} \\
    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn} \\
    &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
    &A \twoheadrightarrow! \; B &&\coloneqq \Sigma (f : A \rightarrow B) , \text{sp-surj}(f) \label{sp-surj-arrow-eqn}
  \end{alignat}
\end{definition}
Over sets, the surjections and split surjections are the same thing, but there
is a difference one we involve non-set types like the circle.

@@

@@

To specify that a value must exist uniquely in HoTT we can use the concept of a
\emph{contraction}.
\begin{definition}[Unique Membership] \label{uniq-memb-def}
  Unique list membership is defined in terms of list membership: it is a
  contraction of it.
  \begin{equation}
    x \inunique \mathit{xs} \coloneqq \text{isContr}(x \in \mathit{xs})
  \end{equation}
\end{definition}
With this we can define manifest Bishop finiteness:
\begin{definition}[Manifest Bishop Finiteness]  \label{bish-def}
  A type is manifest Bishop finite if there exists a list which contains each
  value in the type once.
  \begin{equation}
    \mathcal{B}(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
  \end{equation}
\end{definition}
The only difference between manifest Bishop finiteness and split enumerability
is the membership term: here we require unique membership (\(\inunique\)),
rather than simple membership (\(\in\)).

@@

We have now proved that every manifestly Bishop finite type is split enumerable,
and vice versa.
While the types are not \emph{equivalent} (there are more split enumerable
proofs than there are manifest Bishop finite proofs), they are of equal power,
so any closure proof we have on one can be transferred to the other.
In particular, it means that manifest Bishop finiteness is closed under
\(\Sigma\).
\subsection{From Manifest Bishop Finiteness to Equivalence}
We have seen that split enumerability was in fact a split-surjection in
disguise.
We will now see that manifest Bishop finiteness is in fact an \emph{equivalence}
in disguise.
\emph{equivalence} with \(\mathbf{Fin}\).
\begin{lemma} \label{bishop-equiv}
  Manifest bishop finiteness is equivalent to an equivalence to a finite prefix
  of the natural numbers.
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{lemma}
\begin{proof}
  \begin{align*}
     \mathcal{B}(A) &
    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    && \text{def.~\ref{bish-def} }(\mathcal{B})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(x \in \mathit{xs})
    && \text{eqn.~\ref{uniq-memb-def} } (\inunique)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(\text{fib}_{\text{snd}(\mathit{xs})}(x))
    && \text{eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq A )
    && \text{eqn.~\ref{equiv-def} } (\simeq) \; \qedhere
  \end{align*}
\end{proof}
This proof is almost identical\footnotemark to the proof for
lemma~\ref{split-enum-is-split-surj}: it reveals that
enumeration-based finiteness predicates are simply another perspective on
relation-based ones.
\footnotetext{
  Unfortunately in our formalisation this proof cannot be a single line: for
  performance reasons \(\simeq\) is defined as a record type with eta-equality
  disabled, instead of the definition here which uses \(\Sigma\).
}

@@

CuTT is not so restricted.
Since we have things like function extensionality and transport, we can indeed
prove the finiteness of function types.
Our proof here makes use directly of the univalence axiom, and makes use
furthermore of all the previous closure proofs.
We will prove this closure on split enumerability, rather than on manifest
Bishop finiteness, as it requires slightly less legwork in the proof itself, but
of course we can derive the proof on manifest Bishop finiteness in a few lines.
\begin{lemma} \label{split-enum-pi}
  Split enumerability is closed under dependent functions.
  (\(\Pi\)-types).
  \begin{equation}
    \frac{
      \mathcal{E}!(A) \; \; \; \Pi {(x : A)} , \mathcal{E}!\left( U(x) \right)
    }{
      \mathcal{E}!\left(\Pi {(x : A)} , U(x)\right)
    }
  \end{equation}
\end{lemma}
\begin{proof}
  Let \(A\) be a split enumerable type, and \(U\) be a type family from \(A\),
  which is split enumerable over all points of \(A\).

@@

To remedy the problem, we will use propositional truncation
(def.~\ref{prop-trunc}).
\begin{definition}[Cardinal Finiteness]
  A type \(A\) is cardinally finite if there exists a propositionally truncated
  proof that \(A\) is manifest Bishop finite or equivalent to a finite prefix of
  the natural numbers.
  \begin{equation}
    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq A) \rVert
  \end{equation}
\end{definition}
At first glance, it might seem that we lose any useful properties we could
derive from \(\mathcal{B}\).
Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.
\subsection{Deriving Uniquely-Determined Quantities}
The following two lemmas are proven in
\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
respectively), in much the same way as we have done here.
Our contribution for this section is simply the formalisation.
\begin{lemma}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
  the same cardinality.
  \begin{equation}
    \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
  \end{equation}
\end{lemma}
\begin{proof}
  Let \(A\) be a cardinally-finite type, with proof \(F : \mathcal{C}(A)\).
  Our task is to extract a natural number \(n : \mathbb{N}\) representing the
  cardinality of \(A\), and a propositionally-truncated proof that \(A\) is
  equivalent to \(\mathbf{Fin}(n)\).

@@

\begin{definition}[Manifest Enumerability]
  Manifest enumerability is an enumeration predicate like Bishop finiteness or
  split enumerability with the only difference being a propositionally truncated
  membership proof.
    \mathcal{E}(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert

@@

As a function-based definition, this predicate represents surjections.

>>>>

  Moreover, the inductive type and the container-based type are equivalent, as
  is proven in our formalisation.
\end{definition}
\begin{definition}[Containers] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a pair
  \(S , P\) where \(S\) is a type, the elements of which are called
  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
  the elements of \(P(s)\) are called the \emph{positions} of a container.
  We ``interpret'' a container into a functor defined like so:
  \begin{equation} \label{container-interp}
    \llbracket S , P \rrbracket(A) \coloneqq \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
  \end{equation}
  Membership of a container can be defined like so:
  \begin{equation} \label{container-membership}
    x \in \mathit{xs} \coloneqq \text{fib}_{\text{snd}(\mathit{xs})}(x)
  \end{equation}
\end{definition}

\begin{definition}[Fibers] \label{fibers}
  A fiber \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
  A \rightarrow B\).
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
  \end{agdalisting}
\end{definition}

\begin{definition}[Disjoint Union]
  We define disjoint union as an inductive type.

  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
  \end{agdalisting}

  It is also expressible with only \(\Sigma\):

  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
  \end{agdalisting}

  Although the inductive type definition is slightly more ergonomic.
\end{definition}
\begin{definition}[\(\mathbf{Fin}\)] \label{Fin}
  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\). We
  define it the standard way:
  \begin{equation}
    \begin{alignedat}{3}
      & \mathbf{Fin}(0)     && \coloneqq \bot ; \\
      & \mathbf{Fin}(n + 1) && \coloneqq \top \uplus \mathbf{Fin}(n) ;
    \end{alignedat}
  \end{equation} 
\end{definition}
\subsection{Instances}
Now that we have a suitable definition of finiteness, we will next prove that
some things are finite.
\begin{lemma}
  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are split enumerable.
\end{lemma}
\begin{proof}
  These three types are represented by fairly simple finite sets:
  \begin{align}
    \bot &= \left\{  \right\} \\
    \top &= \left\{ \text{tt}  \right\} \\
    \mathbf{Bool} &= \left\{  \text{false}, \text{true} \right\}
  \end{align}
  Which leads to quite simple proofs of finiteness.

@@

First, we will need to define our terms: in HoTT, surjections are a little more
complex than what you'd find in either MLTT or classical mathematics.
\begin{definition}[Split Surjections] \label{split-surjections}
  We define \emph{split} surjections here \cite[definition
  4.6.1]{hottbook}.
  \begin{alignat}{3}
    &\text{sp-surj}(f)          &&\coloneqq \Pi(y : B) , \text{fib}_f(y) \label{sp-surj-eqn} \\
    &A \twoheadrightarrow! \; B &&\coloneqq \Sigma (f : A \rightarrow B) , \text{sp-surj}(f) \label{sp-surj-arrow-eqn}
  \end{alignat}
\end{definition}
Over sets, the surjections and split surjections are the same thing, but there
is a difference one we involve non-set types like the circle.

@@

\begin{definition}[Decidable Types]
  \begin{equation}
    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
  \end{equation}
\end{definition}
\begin{definition}[Discrete Types]
  A discrete type is one with decidable equality.
  \begin{equation}
    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
  \end{equation}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
  type is a set.
\end{definition}

@@

To specify that a value must exist uniquely in HoTT we can use the concept of a
\emph{contraction}.

\begin{definition}[Homotopy Levels] \label{homotopy-types}
  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
  Some have higher homotopies (paths which aren't unique).
  We actually have a hierarchy of complexity of structure of path spaces in
  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
  \cite[definition 3.1.1]{hottbook}.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{hlevels}
  \end{agdalisting}
\end{definition}

\begin{definition}[Unique Membership] \label{uniq-memb-def}
  Unique list membership is defined in terms of list membership: it is a
  contraction of it.
  \begin{equation}
    x \inunique \mathit{xs} \coloneqq \text{isContr}(x \in \mathit{xs})
  \end{equation}
\end{definition}
With this we can define manifest Bishop finiteness:
\begin{definition}[Manifest Bishop Finiteness]  \label{bish-def}
  A type is manifest Bishop finite if there exists a list which contains each
  value in the type once.
  \begin{equation}
    \mathcal{B}(A) \coloneqq \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \inunique xs
  \end{equation}
\end{definition}
The only difference between manifest Bishop finiteness and split enumerability
is the membership term: here we require unique membership (\(\inunique\)),
rather than simple membership (\(\in\)).

@@

We have now proved that every manifestly Bishop finite type is split enumerable,
and vice versa.
While the types are not \emph{equivalent} (there are more split enumerable
proofs than there are manifest Bishop finite proofs), they are of equal power,
so any closure proof we have on one can be transferred to the other.
In particular, it means that manifest Bishop finiteness is closed under
\(\Sigma\).
\subsection{From Manifest Bishop Finiteness to Equivalence}
We have seen that split enumerability was in fact a split-surjection in
disguise.
We will now see that manifest Bishop finiteness is in fact an \emph{equivalence}
in disguise.
\emph{equivalence} with \(\mathbf{Fin}\).

\begin{definition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{alignat}{2}
    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f) \label{equiv-def}
  \end{alignat}
\end{definition}
\begin{lemma} \label{bishop-equiv}
  Manifest bishop finiteness is equivalent to an equivalence to a finite prefix
  of the natural numbers.
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{lemma}
\begin{proof}
  \begin{align*}
     \mathcal{B}(A) &
    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    && \text{def.~\ref{bish-def} }(\mathcal{B})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(x \in \mathit{xs})
    && \text{eqn.~\ref{uniq-memb-def} } (\inunique)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(\text{fib}_{\text{snd}(\mathit{xs})}(x))
    && \text{eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq A )
    && \text{eqn.~\ref{equiv-def} } (\simeq) \; \qedhere
  \end{align*}
\end{proof}
This proof is almost identical\footnotemark to the proof for
lemma~\ref{split-enum-is-split-surj}: it reveals that
enumeration-based finiteness predicates are simply another perspective on
relation-based ones.
\footnotetext{
  Unfortunately in our formalisation this proof cannot be a single line: for
  performance reasons \(\simeq\) is defined as a record type with eta-equality
  disabled, instead of the definition here which uses \(\Sigma\).
}

@@

CuTT is not so restricted.
Since we have things like function extensionality and transport, we can indeed
prove the finiteness of function types.
Our proof here makes use directly of the univalence axiom, and makes use
furthermore of all the previous closure proofs.
We will prove this closure on split enumerability, rather than on manifest
Bishop finiteness, as it requires slightly less legwork in the proof itself, but
of course we can derive the proof on manifest Bishop finiteness in a few lines.
\begin{theorem} \label{split-enum-pi}
  Split enumerability is closed under dependent functions.
  (\(\Pi\)-types).
  \begin{equation}
    \frac{
      \mathcal{E}!(A) \; \; \; \Pi {(x : A)} , \mathcal{E}!\left( U(x) \right)
    }{
      \mathcal{E}!\left(\Pi {(x : A)} , U(x)\right)
    }
  \end{equation}
\end{theorem}
\begin{proof}
  Let \(A\) be a split enumerable type, and \(U\) be a type family from \(A\),
  which is split enumerable over all points of \(A\).

@@

To remedy the problem, we will use propositional truncation
(def.~\ref{prop-trunc}).

\begin{definition}[Higher Inductive Types] \label{HITs}
  Normal inductive types have \emph{point} constructors: constructors which
  construct values of the type.
  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
  construct paths in the type.
\end{definition}
\begin{definition}[Propositional Truncation] \label{prop-trunc}
  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
  proof of \(A\) \cite[3.7]{hottbook}.
  In other words, it is a proof that some \(A\) exists, but it does not tell you
  \emph{which} \(A\).

  It is defined as a Higher Inductive Type:
  \begin{equation} {
    \begin{alignedat}{3}
      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &:& \; A \rightarrow \lVert A \rVert ; \\
                              | & \; \text{squash}     &:& \; \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
    \end{alignedat} }
  \end{equation}
  We will use two eliminators from \(\lVert A \rVert\) in this paper.
  \begin{enumerate}
  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
  \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
    function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
    choice of \(A\):
    \[\Pi {(x , y : A)} , f(x) \equiv f(y) \]
    Formally speaking, \(f\) needs to be ``coherently constant''
    \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
    \(n\)-type for some finite \(n\).
  \end{enumerate}
\end{definition}
\begin{definition}[Cardinal Finiteness]
  A type \(A\) is cardinally finite if there exists a propositionally truncated
  proof that \(A\) is manifest Bishop finite or equivalent to a finite prefix of
  the natural numbers.
  \begin{equation}
    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq A) \rVert
  \end{equation}
\end{definition}
At first glance, it might seem that we lose any useful properties we could
derive from \(\mathcal{B}\).
Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.
\subsection{Deriving Uniquely-Determined Quantities}
The following two lemmas are proven in
\cite{yorgeyCombinatorialSpeciesLabelled2014} (Proposition 2.4.9 and 2.4.10,
respectively), in much the same way as we have done here.
Our contribution for this section is simply the formalisation.
\begin{lemma}
  Given a cardinally finite type, we can derive the type's cardinality, as well
  as a propositionally truncated proof of equivalence with \(\textbf{Fin}\)s of
  the same cardinality.
  \begin{equation}
    \mathcal{C}(A) \rightarrow \Sigma {(n : \mathbb{N})} , \lVert \textbf{Fin}(n) \simeq A \rVert
  \end{equation}
\end{lemma}
\begin{proof}
  Let \(A\) be a cardinally-finite type, with proof \(F : \mathcal{C}(A)\).
  Our task is to extract a natural number \(n : \mathbb{N}\) representing the
  cardinality of \(A\), and a propositionally-truncated proof that \(A\) is
  equivalent to \(\mathbf{Fin}(n)\).

@@

\begin{definition}[Manifest Enumerability]
  Manifest enumerability is an enumeration predicate like Bishop finiteness or
  split enumerability with the only difference being a propositionally truncated
  membership proof.
  \begin{equation}
    \mathcal{E}(A) \coloneqq \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
\end{definition}
As a function-based definition, this predicate represents surjections.

\begin{definition}[Surjections] \label{surjections}
  We define proper surjections (not split surjections) here \cite[definition
  4.6.1]{hottbook}.
  \begin{alignat}{3}
    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \label{surj-eqn} \\
    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn}
  \end{alignat}

@@

### 2020-06-17 sections/introduction.tex
moved out-directory instructions to makefile
moved definitions to their place of first usage

\begin{definition}[Disjoint Union]
  We define disjoint union as an inductive type.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
  \end{agdalisting}
  It is also expressible with only \(\Sigma\):
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
  \end{agdalisting}
\end{definition}
\begin{definition}[Equalities, equivalences, and paths]
  \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
  Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
  so the distinction isn't terribly important in usage: we will only use one
  or the other as a suggestion of how we constructed it or how it is to be
  used.
\end{definition}

@@

\footnotetext{
  Actually, CuTT does have an identity type with similar semantics to the
  identity type in MLTT.
  We do not use this type anywhere in our work, however, so we will not consider
  it here.
}
\begin{definition}[Homotopy Levels] \label{homotopy-types}
  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
  Some have higher homotopies (paths which aren't unique).
  We actually have a hierarchy of complexity of structure of path spaces in
  types, starting with the contractions \cite[definition 3.11.1]{hottbook}, then
  the mere propositions \cite[definition 3.3.1]{hottbook}, and the sets
  \cite[definition 3.1.1]{hottbook}.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{hlevels}
  \end{agdalisting}
\end{definition}

\begin{definition}[Fibers] \label{fibers}
  A fiber \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
  A \rightarrow B\).
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
  \end{agdalisting}
\end{definition}
\begin{definition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{alignat}{2}
    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f) \label{equiv-def}
  \end{alignat}
\end{definition}
\begin{definition}[Decidable Types]
  \begin{equation}
    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
  \end{equation}
\end{definition}
\begin{definition}[Discrete Types]
  A discrete type is one with decidable equality.
  \begin{equation}
    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
  \end{equation}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
  type is a set.
\end{definition}
\begin{definition}[Higher Inductive Types] \label{HITs}
  Normal inductive types have \emph{point} constructors: constructors which
  construct values of the type.
  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
  construct paths in the type.
\end{definition}
\begin{definition}[Propositional Truncation] \label{prop-trunc}
  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
  proof of \(A\) \cite[3.7]{hottbook}.
  In other words, it is a proof that some \(A\) exists, but it does not tell you
  \emph{which} \(A\).

  It is defined as a Higher Inductive Type:
  \begin{equation} {
    \begin{alignedat}{3}
      \lVert A \rVert \coloneqq & \; \lvert \wc \rvert &:& \; A \rightarrow \lVert A \rVert ; \\
                              | & \; \text{squash}     &:& \; \Pi {(x, y : \lVert A \rVert)} , x \equiv y  ; 
    \end{alignedat} }
  \end{equation}
  We will use two eliminators from \(\lVert A \rVert\) in this paper.
  \begin{enumerate}
  \item \label{elim-prop-prop} For any function \(A \rightarrow B\), where
    \(\text{isProp}(B)\), we have a function \(\lVert A \rVert \rightarrow B\).
  \item \label{elim-prop-coh} We can eliminate from \(\lVert A \rVert\) with a
    function \(f : A \rightarrow B\) iff \(f\) ``doesn't care'' about the
    choice of \(A\):
    \[\Pi {(x , y : A)} , f(x) \equiv f(y) \]
    Formally speaking, \(f\) needs to be ``coherently constant''
    \cite{krausGeneralUniversalProperty2015}, and \(B\) needs to be an
    \(n\)-type for some finite \(n\).
  \end{enumerate}
\end{definition}

>>>>

@@

\footnotetext{
  Actually, CuTT does have an identity type with similar semantics to the
  identity type in MLTT.
  We do not use this type anywhere in our work, however, so we will not consider
  it here.
}
